Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-31 code-tests.tex
Added separate file for FLD test problem
Sedov blast test added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sedov Explosion}
\label{sec.tests.sedov}
\red{(Elizabeth)}
Problem type 7.  One of Alexei's test problems.  
Also problem 7.4 in the FLASH method paper.
This tests the hydro in the strong-shock limit.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Representative FLD test}
\label{sec.tests.fld}
\red{(Dan)}
One test for the implicit FLD: best/hardest one.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sedov Explosion}
\label{sec.tests.sedov}
\red{(Elizabeth)}

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
\caption{Density slice from the Sedov Blast Test at $t = 0.07$. Left-hand image shows the results from PPM ({\tt HydroMethod = 0}), right-hand image shows the results from Zeus ({\tt Hydromethod = 2}). Notably, the Zeus shock front has progressed less far than forthe PPM run. This is due to energy loss when conserving internal, not total, energy.}
\label{fig.sedov1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{}
\label{fig.sedov2}
\end{center}
\end{figure}

The Sedov Blast Test \citep{Sedov1959} models an intense explosion, initiated by depositing thermal energy into a homogenous distribution of gas. The result is a strong spherical shock wave centered on the point of energy injection.  This problem is a popular test of astrophysical numerical codes for three reasons: Firstly, it is particularly appropriate to astrophysics since it represents the situation when a supernovae explosion occurs. Secondly, it has an exact analytical solution whereby the shock front's radial position is given by:

\begin{equation}
r(t) = \left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
\end{equation}

\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the background density and $\alpha = 1.0$ for cylindrical symmetry and an ideal gas with $\gamma = 1.4$. For the full derivation see \citet{Sedov1959}. This solution makes it possible to assess how well the code performs. Thirdly, the spherical shock is a challenging problem for numerical codes (both grid and particle) since the shock wave increases in size as the simulation progresses and its symmetrical nature highlights any directional preferences that grid codes can succumb to. The test presented here is the two-dimensional version that is included in the Enzo distribution. The three-dimensional results from this test, both for Enzo and three other leading astrophysics codes, can be founds in \citet{Tasker2008}.

In the initial state, the box contains a homogenous distribution of gas at a density of 1 (note that, in the absense of gravity, this problem is scale-free and so without units). Thermal energy is deposited into a single cell at the center of the box with $E_0 = 10.0$. The problem is completed in two dimensions with reflecting boundary conditions (the default for Enzo) and uses a boxsize of side $1$. For this problem, we selected a top grid of $100 \times 100$ and four levels of refinement, placed based on shock location and the slope in density and total energy (this corresponds to cell flagging methods 1 and 3 in the Enzo parameter file). The exception to this scheme is in the initial conditions where grids are placed directly around the injection point. The results were assessed at $t = 0.07$, which corresponding to a time just before the shock reaches the box edge (see Figure~\ref{fig.sedov1}).

Figure~\ref{fig.sedov2} shows the radial profiles for the simulation run with the PPM hydro-solver (blue dashed line) and the Zeus hydro-solver (red dash-dot-dot line) together with the analytical solution (black solid line). Clockwise from the top-left are density, velocity, internal energy and pressure. PPM matches the analytical solution extremely well for all quantities. However, Zeus' shock front lags behind the analytical position. This can also be seen in the slices shown in Figure~\ref{fig.sedov1}. The cause of this discrepancy is that Zeus shows a substantial energy lost during the first few timesteps and produces a diamond-shaped, rather than spherical, shockfront during this time. After this, the code correctly conserves energy but this intial energy loss remains clearly visible in the position of the shock at $t = 0.07$. This problem was addressed directly by \citet{Clarke2010}, who attributed the source of the issue to this version of Zeus solving the internal, rather than total, energy equation. In situations with strong energy gradients, this choice caused an energy loss and the artificial viscosity produces the direction-dependent shockfront shape. In their paper, \citet{Clarke2010} presents results from an alternative version of Zeus that conserves total energy. This problem is less marked for smaller energy gradients and it should be noted that Zeus' stability and speed make it a highly competitive choice, despite the disagreements in this test.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{code-tests-fld}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Representative thermal conduction test}
\label{sec.tests.conduct}
\red{(Brian)}
One test for the thermal conduction (maybe 1 for isotropic, 1 for
anisotropic).
